High-level cognition is supported by at least second order dynamic correlations in neural activity patterns

### 2019-08-21 paper/main.tex

We validated our approach using synthetic data where the underlying
correlations were known.  We then applied our framework to a
neuroimaging dataset collected as 125 participants listened to either
an audio recording of a ten-minute story or a temporally scrambled
version of the story, or underwent a resting state
scan~\citep{SimoEtal16}.  We used a subset of the data to train
across-participant classifiers to decode listening times using a blend
of neural features (comprising neural activity patterns, as well as
different orders of correlations between those patterns that were
inferred using our computational framework).  We found that both the
PCA-based and eigenvector centrality-based approaches yielded neural
patterns that could be used to decode accurately.  Both approaches
also yielded the best decoding accuracy for data collected during
(intact) story listening when high-order (PCA:
second-order; eigenvector centrality: fourth-order) dynamic
correlation patterns were included as features.  When we trained
classifiers on the scambled stories or resting state data, only
lower-order dynamic patterns were informative to the decoders.  Taken
together, our results indicate that high-level cognition is supported
by high-order dynamic patterns of communication between brain
structures.  \textbf{JRM STOPPED HERE}

\section*{Methods}
A major challenge to studying such patterns is that typically neither
the correlations nor the hierarchical organizations of those
correlations may be directly observed.  Rather, these fundamental
properties must be inferred indirectly by examining the observable
parts of the system-- e.g., the behaviors of the individual units of
that system.  Here we propose a series of mathematical operations that may be used to approximate dynamic correlations at a range of scales (i.e., orders of interaction).

>>>>

We validated our approach using synthetic data where the underlying
correlations were known.  We then applied our framework to a
neuroimaging dataset collected as 125 participants listened to either
an audio recording of a ten-minute story or a temporally scrambled
version of the story, or underwent a resting state
scan~\citep{SimoEtal16}.  We used a subset of the data to train
across-participant classifiers to decode listening times using a blend
of neural features (comprising neural activity patterns, as well as
different orders of correlations between those patterns that were
inferred using our computational framework).  We found that both the
PCA-based and eigenvector centrality-based approaches yielded neural
patterns that could be used to decode accurately.  Both approaches
also yielded the best decoding accuracy for data collected during
(intact) story listening when high-order (PCA:
second-order; eigenvector centrality: fourth-order) dynamic
correlation patterns were included as features.  When we trained
classifiers on the scambled stories or resting state data, only
lower-order dynamic patterns were informative to the decoders.  Taken
together, our results indicate that high-level cognition is supported
by high-order dynamic patterns of communication between brain
structures.

\section*{Methods}
Our general approach to comprises four general steps.  First, we
derive a kernel-based approach to computing dynamic pairwise
correlations in a multivariate timeseries.  Next, we apply a
dimensionality reduction step to \textbf{JRM STOPPED HERE}

A major challenge to studying such patterns is that typically neither
the correlations nor the hierarchical organizations of those
correlations may be directly observed.  Rather, these fundamental
properties must be inferred indirectly by examining the observable
parts of the system-- e.g., the behaviors of the individual units of
that system.  Here we propose a series of mathematical operations that may be used to approximate dynamic correlations at a range of scales (i.e., orders of interaction).
